Benchmark!!1

### 2021-02-11 paper/main.tex
progrss on the paper

\subsection{Dynamic programming for column generation}
Many state-of-the-art exact methods for vehicle routing problems rely
on column generation (see e.g.~\cite{Baldacci:2012survey}). In such
methods, a significant amount of CPU effort is spent solving the
\emph{pricing subproblem}, which differs depending on the problem at
hand but is nonetheless often an elementary shortest path problem with
resource constraints (ESPPRC). While it can be solved heuristically,
in order to establish optimality it eventually needs to be solved
exactly as well. For this benchmark we implement a dynamic
programming algorithm similar to the one by~\cite{feillet:2004exact}
to solve the ESPPRC, albeit without time windows. One difficulty in
this problem stems from the existence of negative-cost cycles in the
graph.

In order to simulate an environment similar to the one encountered in
column generation for VRPs, we derive an ESPPRC instance from each TSP
solution by using an auxilliary distance matrix $c'$ from the
distance matrix $c$, as follows:
\begin{enumerate}
\item For each vertex $j$, let $\pi_j$ be the cost of arc $(i, j)$,
  where $i$ is the predecessor of $j$ in the TSP solution.
\item For each arc $(i, j)$, set $c'_{ij}$ to $c_{ij} - d_j$.
\end{enumerate}
Additionally, we generate resource consumption as follows: we consider
6 resources numbered from 0 to 5, and any vertex $i$ consumes resource
$r$ iff the $r^{th}$ bit of $i$'s binary representation is 1
(considering that the first bit is bit 0). That is, in C syntax, iff
$i & (1 << r) > 0$. Capacity is set to 1 for each resource.
The algorithm then computes the elementary shortest path with resource
constraints on $c'$ from 0 to 0 (1 to 1 in Julia).

@@

\subsection{Other general cross-language considerations}
Subjective considerations:
\begin{itemize}
\item Typically easier to program using Python, Julia or JavaScript.
\item Typically harder to program using Rust than anything else but
  might be due to lack of experience. However my experience in
  JavaScript is much lower than in C++ and I still did everything
  faster in JavaScript than in C++.
\end{itemize}

@@

\subsection{Numba}
Numba is not yet feature-complete for Numpy. A direct consequence is that some
benchmarks cannot be implemented for Numba. For example
\lstinline{numpy.insert()} is not implemented, but this function is
needed for the $LN$ benchmark. Therefore the $LNS$ benchmark cannot be
implemented with Numba. Additionally, we were not able to implement
the ESPPRC benchmark in a satisfying manner. In our experience, Numba
error messages are sometimes unrelated to the issue causing them, or
too obscure to make sense, keeping in mind the premises of this
paper. It would certainly be possible to implement a dynamic
programming algorithm for the ESPPRC benchmark using Numba's JIT
compilationfeatures, however our conclusion is that it would not be
possible to do so without expert knowledge of Numba.

@@

\subsection{Java}
Mention inconsistencies between Java versions: code written for Java 8
does not compile with Java 11. Because Pair is not in the API any
more.

\subsection{Rust}
Issues:
\begin{itemize}
\item Changing from i32 to i64 is problematic somehow
\item Not possible to have refs to both successors and predecessor in
  Label, hence had to handle indices by hand in specific espprc implementation
\end{itemize}

\section{Experiments}
All programs are run concurrently on an AMD Ryzen 5 3600 6-Core
Processor at 3.6 GHz with 16 GB RAM running Linux. Twelve threads can
run in parallel, however in our setting at most 5 threads are used at
any given time.

@@

>>>>

\subsection{Dynamic programming for column generation}
\label{sec:espprc}
Many state-of-the-art exact methods for vehicle routing problems rely
on column generation (see e.g.~\cite{Baldacci:2012survey}). In such
methods, a significant amount of CPU effort is spent solving the
\emph{pricing subproblem}, which differs depending on the problem at
hand but is nonetheless often an elementary shortest path problem with
resource constraints (ESPPRC). While it can be solved heuristically,
in order to establish optimality it eventually needs to be solved
exactly as well. For this benchmark we implement a dynamic
programming algorithm similar to the one by~\cite{feillet:2004exact}
to solve the ESPPRC, albeit without time windows. One difficulty in
this problem stems from the existence of negative-cost cycles in the
graph.

In order to simulate an environment similar to the one encountered in
column generation for VRPs, we derive an ESPPRC instance from each TSP
solution by using an auxilliary distance matrix $c'$ from the
distance matrix $c$, as follows:
\begin{enumerate}
\item For each vertex $j$, let $\pi_j$ be the cost of arc $(i, j)$,
  where $i$ is the predecessor of $j$ in the TSP solution.
\item For each arc $(i, j)$, set $c'_{ij}$ to $c_{ij} - d_j$.
\end{enumerate}
Additionally, we generate resource consumption as follows: we consider
6 resources numbered from 0 to 5, and any vertex $i$ consumes resource
$r$ iff the $r^{th}$ bit of $i$'s binary representation is 1
(considering that the first bit is bit 0). That is, in C syntax, iff
\verb|i & (1 << r) > 0|. Capacity is set to 1 for each resource.
The algorithm then computes the elementary shortest path with resource
constraints on $c'$ from 0 to 0 (1 to 1 in Julia).

@@

\subsection{Other general cross-language considerations}
Here we discuss considerations that are subjective in nature, as they
relate to our experience implementing all benchmarks in all
languages. We need to take into account our experience using various
languages prior to the beginning of this work:
\begin{itemize}
\item C++: numerous projects over 20 years.
\item Python: numerous projects over 15 years; no experience of Numpy
  or Numba.
\item Java: PhD work 15 years ago, one minor project 5 years ago.
\item Julia: one ongoing research project started 3 years ago.
\item Javascript: a few hobby projects 5-7 years ago.
\item Rust: no experience at all.
\end{itemize}
To add to the subjectivity, each benchmark was first implemented in
Python then in other languages, with Rust typically being the last one
to be implemented. Therefore the languages implemented last were
supposedly easier, due to increased familiarity with the
algorithm. For all reasons mentioned above, measuring the
time spent implementing each benchmark in each language would be
meaningless, therefore we limit ourselves to subjective opinion here.
Keeping this in mind, we see a subset of perceived
\emph{easy} languages, as in ``easy to program in and get things running
with the expected outcome'', which consists of Python, Julia and
JavaScript. These 
easy languages typically allow to program the same algorithms faster
than the other ones with the same results, performance considerations
aside.
On the other hand, we found the process of getting benchmarks to work
with Rust to be much more tedious than with any other language,
although this is certainly due to lack of experience. Nonetheless, the
JavaScript implementation was typically easier than the C++ or Java ones
despite having more experience with C++ and Java, so
experience does not explain everything.
Ultimately, there is little doubt that any experienced programmer in
any of these languages would have no trouble implementing any of these
benchmarks. Nonetheless, this programmer with experience in mostly C++
and Python implemented every benchmark faster in Python, JavaScript or Julia
than in C++.

We also want to underline that no matter the language, it is possible
to write both efficient and inefficient code. Familiarity and
experience are beneficial; beginners in a language are more likely to
write inefficient code. This being said, not every language is equal
when it comes to documenting how to write efficient code. We note that
the official Julia documentation has a whole section dedicated to that
topic~\cite{julia-performance}. There also exists an automated tool to
detect inefficient code patterns, based on the official
documentation~\cite{julia-traceur}.

@@

\subsection{Numba}
Numba is not yet feature-complete for Numpy. A direct consequence is that some
benchmarks cannot be implemented for Numba. For example
\verb|numpy.insert()| is not implemented, but this function is
needed for the $LNS$ benchmark, therefore the $LNS$ benchmark cannot be
implemented with Numba. Additionally, we were not able to implement
the ESPPRC benchmark in a satisfying manner. In our experience, Numba
error messages are sometimes unrelated to the issue causing them, or
too obscure to make sense, keeping in mind the premises of this
paper. It would certainly be possible to implement a dynamic
programming algorithm for the ESPPRC benchmark using Numba's JIT
compilation features, however our conclusion is that it would not be
possible to do so without advanced knowledge of Numba.

@@

\subsection{Java}
The ecosystem of Java implementations and versions can be hard to
navigate. There exist multiple compilers, multiple virtual
machines, and there is a new Java version every 6 months, sometimes
breaking backwards compatibility. The official reference
implementation is OpenJDK since version 7~\cite{openjdk-reference}.
We use OpenJDK version 11, which is the current long-term support version as
of writing this article.

\subsection{Rust}
As mentioned in Section~\ref{sec:espprc}, Rust does not allow a given
\verb|struct| to have multiple references to the same \verb|struct|
type. For this reason, we needed to develop a different implementation
of the espprc benchmark. In general Rust puts restrictions on what
constitutes valid code, for the sake of safety. This is a valuable
feature in systems programming, which is the main application that Rust was
designed for. This also means that writing code that compiles and runs
is actually a more tedious task, but that once the code compiles,
there is also a higher chance that it will run according to
expectations, like with the Ada programming language for instance.

In the context of this study, these benefits did not appear relevant
however: each benchmark was programmed in all other languages as well,
providing the same results as with the Rust implementation, while no
evidence of memory leak was observed. As mentioned above, The Rust
implementation was typically way more time-consuming than other
implementations, although one reason for that is lack of
experience. Nonetheless, in the context of this study, the generally
assumed benefits of using Rust are not apparent, with the notable
exception of good runtime performance.

\section{Experiments}
All programs are run concurrently on an AMD Ryzen 5 3600 6-Core
Processor at 3.6 GHz with 16 GB RAM running Linux. Twelve threads can
run in parallel, however in our setting at most 5 threads are used at
any given time. For every benchmark and every instance, CPU times are
reported as a ratio of a \emph{reference time} for that benchmark and
instance. In practice, for each benchmark and instance, the reference
time is the time required by the c++98 implementation for that
benchmark and that instance. Therefore the c++98 implementation always
has a CPU ratio of 1.

@@

\subsection{Cross-platform comparison}
We now look at the same benchmarks, using the same languages, but running
in a different environment. We compare runs on an Intel i5-8265U
4-core CPU at 1.6 GHz and on the previously mentioned AMD CPU. The
goal is not to observe which CPU provides the best performance, rather
to determine whether the experimental observations made with one CPU
are valid with the other. For this reason, we stick to reporting CPU
times as a ratio to the CPU time required by the c++98 implementation.

Perhaps also run on WU cluster and observe differences.
